---
title: "Documentation of Dev-concepts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{documentation-dev-concepts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Github-actions workflow for shinyapps.io

The aim is to stup a configuration that customizes Github actions in a way that
allows to deploy the shiny App to shinyapps.io automatically. Specifically:

1. The main branch reflects the App in production state
2. Any PR reflects various tests i.e. different dev-states of the App
3. Once a PR gets merged into main, the production App is updates with the 
PR-feature and automatically re-deployed
4. The testing/dev-App versions are also automatically deployed to hidden links
5. Eventually, a testing workflow a-la cypress can be employed onto the 
different test/dev-App versions

### Useful notes

- https://blog.rmhogervorst.nl/blog/2021/02/27/deploy-to-shinyapps-io-from-github-actions/

- https://www.r-bloggers.com/2021/02/deploy-to-shinyapps-io-from-github-actions/


### Extracting `shiny.io` secrets and defining useful environment variables

1.Via a project/package local `.Renviron` with the following content:
```{bash, eval = FALSE}
# R_LIBS_USER=~/Dropbox/libraries/R/4.0
SHINY_ACC_NAME=cologne-analytics
TOKEN=315FD675C450786E52E88588B0554D83
SECRET=yXqZjbaYUe/oYvwEStsRh7inKQ8o5w21ys5IgbGg
MASTERNAME=TaskSVVdcDB
TESTNAME=TaskSVVdcDBtesting
```

__*Note:*__

The first line needs to be run if the setup is tested locally on a machine. If 
the script is tested (locally) in a docker container this line has to be 
commented out.

2.Via Github secrets/variables either organization or repository wide:

- https://docs.github.com/en/actions/security-guides/encrypted-secrets -> `{{ secrets.mysupersecret }}`
- https://docs.github.com/en/actions/learn-github-actions/variables -> `{{ vars.myvars }}`

### Setting up a deployment script

Use the following `R`-template inside directly `./` e.g. `deploy.R` at the 
`app.R` level:
```{r, eval = FALSE}
# A function to stop the script when one of the variables cannot be found.
# Also to strip quotation marks from the secrets when you supplied them
# (maybe it is just easier to never use them).
error_on_missing_name <- function(name) {
  var <- Sys.getenv(name, unset = NA)
  if (is.na(var)) {
    stop(paste0("cannot find ", name, " !"), call. = FALSE)
  }
  gsub("\"", "", var)
}
# Authenticate
rsconnect::setAccountInfo(
  name = error_on_missing_name("SHINY_ACC_NAME"),
  token = error_on_missing_name("TOKEN"),
  secret = error_on_missing_name("SECRET")
)
# Deploy the application.
rsconnect::deployApp(forceUpdate = TRUE)
#   appFiles = c(
#     "app.R" # , you can specify which files to deploy,
#     # or keep this NULL to deploy everything
#   ),
#   appName = error_on_missing_name("MASTERNAME"),
#   appTitle = "shinyapplication"
# )
```

### Dockerfile

Use the following template Dockerfile:

```{bash, eval = FALSE}
FROM rocker/shiny-verse
RUN install2.r rsconnect golem htmltools config writexl shiny.semantic gargoyle
RUN install2.r reactable magrittr plotly auth0 hunspell spelling
RUN installGithub.r TASK-no/TaskAnalyticsTB \
&& rm -rf /tmp/downloaded_packages/
WORKDIR /home/TaskSVVdcDB
COPY app.R app.R
COPY deploy.R deploy.R
COPY DESCRIPTION DESCRIPTION
COPY NAMESPACE NAMESPACE
ADD data data
ADD inst inst
ADD R R
CMD Rscript deploy.R
```

### Testing part I

Source `deploy.R` inside `golem::get_golem_wd()` (environment vars must be read
with first line). Then, check the container setups by running the following in
sudo:
```{bash, eval = FALSE}
IMAGENAME=<NAMETHISSOMETHING>
sudo docker build -t $IMAGENAME .
sudo docker run --env-file .Renviron $IMAGENAME
sudo docker ps
```
Literally, this deploys the App on `shinyapps.io`!

Ease the testing by automation of docker deployment with a `autodeploy.sh`:
```{bash, eval = FALSE}
#!/bin/bash
IMAGENAME=golemautodeploy
sudo docker build -t $IMAGENAME .
sudo docker run --env-file .Renviron $IMAGENAME
sudo docker ps
```
Run `chmdod +x autodeploy.sh` and `./autodeploy.sh`; if repeated regularly one
can even define an alias or a command!


### Create the Github actions profile

Create the directory `.github/workflows` with two files inside:
```{bash, eval = FALSE}
name: Run on push master, main

# Controls when the action will run.
on:
  # Triggers the workflow on push or pull request events but only for the main branch
  push:
    branches: [ main, master ]

  # Allows you to run this workflow manually from the Actions tab
  workflow_dispatch:

# A workflow run is made up of one or more jobs that can run sequentially or in parallel
jobs:
  # This workflow contains a single job called "build"
  build:
    # The type of runner that the job will run on
    runs-on: Ubuntu-20.04

    # Steps represent a sequence of tasks that will be executed as part of the job
    steps:
      # Checks-out your repository under $GITHUB_WORKSPACE, so your job can access it
      - uses: actions/checkout@v2

      # build the docker image and give it the name main
      - name: Build image
        run: docker build -t main .
      # run the docker image supply the secrets from the github secrets store.
      - name: execute
        run: >
          docker run
          -e SHINY_ACC_NAME=${{ vars.SHINY_ACC_NAME }}
          -e TOKEN=${{secrets.TOKEN}}
          -e SECRET=${{secrets.SECRET}}
          -e MASTERNAME=${{vars.MASTER_NAME}}
          main
```
```{bash, eval = FALSE}

name: PR

# Run on all pull requests
on:
  pull_request: {}

  # Allows you to run this workflow manually from the Actions tab
  workflow_dispatch:

# A workflow run is made up of one or more jobs that can run sequentially or in parallel
jobs:
  # This workflow contains a single job called "build"
  build:
    # The type of runner that the job will run on
    runs-on: Ubuntu-20.04

    # Steps represent a sequence of tasks that will be executed as part of the job
    steps:
      # Checks-out your repository under $GITHUB_WORKSPACE, so your job can access it
      - uses: actions/checkout@v2

      # Build image
      - name: Build image
        run: docker build -t pullrequestimage .
      # run the image (but supply the TESTNAME in stead of MASTERNAME, so the
      # app ends up in a different place.
      - name: execute
        run: >
          docker run
          -e SHINY_ACC_NAME=${{vars.SHINY_ACC_NAME}}
          -e TOKEN=${{secrets.TOKEN}}
          -e SECRET=${{secrets.SECRET}}
          -e MASTERNAME=${{vars.TEST_NAME}}
          pullrequestimage

```


## Set up `auth0` for authentication

### Notes/resources

- the R package `https://curso-r.github.io/auth0/`
- the Github issue explaining the workflow with `{golem}` https://github.com/ThinkR-open/golem/issues/327

### In details some of the account settings

...
